---
output: md_document
---

```{r setup}
library(icesTAF)
library(data.tree)

repoName <- "cod.27.47d20"
```

# Creating a TAF analysis

The first step is asking the ICES secratariate to create a repository for your stock.  In this example we will
work with North Sea cod: `r repoName` in 2019 - the repository name for the assessment will be 
`r repoName`.  When a new repository is created on GitHub you should clone the repository to your
own computer.  From here we will assume you have cloned the

The first step in creating a TAF analysis is to set out the basic folder and file structure of the project.
This is done using the function [`taf.skeleton`](https://rdrr.io/cran/icesTAF/man/taf.skeleton.html).
which creates the following structure in your working directory

```{r taf-skeleton-output, results='asis', echo=FALSE}
taf.skeleton(repoName)
paths <- dir(repoName, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(repoName, "/bootstrap/initial/data/"), paths)
paths <- data.frame(pathString = paths)
tree <- as.Node(paths)
res <- c("\`\`\`r",
  capture.output(print(as.data.frame(tree), row.names = FALSE))[-1],
  "\`\`\`")

cat(res, sep = "\n")
```

