---
output: md_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(icesTAF)
library(data.tree)

repoName <- "cod.27.47d20"

if (dir.exists(repoName)) {
  unlink(repoName, recursive = TRUE)
}

formatRcode <- function(x) {
  c("\`\`\`r",
    x,
    "\`\`\`")
}

printDir <- function() {
  # get skeleton path structure
  paths <- dir(repoName, full.names = TRUE, recursive = TRUE)

  # make a data.tree and print it
  tree <- as.Node(data.frame(pathString = paths))
  res <- capture.output(print(as.data.frame(tree), row.names = FALSE))[-1]
  cat(formatRcode(res), sep = "\n")
}

taflink <- function(x) {
  sprintf("[`%1$s`](https://rdrr.io/cran/icesTAF/man/%1$s.html)", x)
}
```

This page was based on using the `icesTAF` package version ```r packageVersion("icesTAF")```
dated ```r packageDate("icesTAF")```.

## In this guide

* [Clone](#clone)
* [Make Skeleton](#make-skeleton)
* [Upload initial data](#upload-initial-data)
  - [Upload files](#upload-files)
  - [Make data available](#make-data-available-to-assessment)

## Clone

The first step is asking the ICES secratariate to create a repository for your stock.  In this example we will
work with North Sea cod: `r repoName` in 2019 - the repository name for the assessment will be 
`r repoName`.  When a new repository is created on GitHub you should clone the repository to your
own computer.  From here we will assume you have cloned the

## Make skeleton

The first step in creating a TAF analysis is to set out the basic folder and file structure of the project.
This is done using the function `r taflink("taf.skeleton")`.
which creates the following structure in your working directory

```{r taf-skeleton-output, results='asis', echo=FALSE}
# make skeleton
taf.skeleton(repoName, force = TRUE)

# get skeleton path structure
paths <- dir(repoName, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(repoName, "/bootstrap/initial/data/"), paths)

# make a data.tree and print it
tree <- as.Node(data.frame(pathString = paths))
res <- capture.output(print(as.data.frame(tree), row.names = FALSE))[-1]
cat(formatRcode(res), sep = "\n")
```


## Upload initial data

The next step is to set up the data requirements for your assessment.  There are 
three ways to get data into an assessment. 1. upload files directly, 2. get a 
file from the web, and 3. use some Rcode to access a webservice.  We will cover 
1. and 2. here.  See https://github.com/ices-taf/doc/wiki/Example-data-records
for more information.

### Upload files

To upload a file called `catch.csv` - the contents might be something like:
```
year, catches
2015, 2109
2016, 3455
2017, 3466
2018, 2050
```

simply copy the file in to the `bootstrap/initial/data` folder.  The directory structure will now look like this:

```{r data-file, results='asis', echo=FALSE}
cat("year, catches
2015, 2109
2016, 3455
2017, 3466
2018, 2050", file = paste0(repoName, "/bootstrap/initial/data/catch.csv"))
printDir()
```

### Make data available to assessment

So far we have uploaded data into the intial data folder, but to make it available to the assessment
it needs to be in the `bootstrap/data` folder, and the only way to do this is to create a file referenceing
the data and then run the function `r taflink("taf.bootstrap")`.
To reference the data we create a file called `DATA.bib` using the helper function `r taflink("draft.data")`, 
and add a few fields:

```{r draft-data, echo=2}
setwd(repoName)
draft.data(originator = "WGNSSK", 
           title = "Catch data for cod.27.347d", 
           period = "2015-2018")
setwd("..")
```

Then we write this to the `DATA.bib` file by specifying the `file` argument:
```{r draft-data2, echo=2}
setwd(repoName)
draft.data(originator = "WGNSSK", 
           title = "Catch data for cod.27.347d", 
           period = "2015-2018",
           file = "bootstrap/DATA.bib")
setwd("..")
```

Finally we run `taf.bootstrap()` to get the data into the `bootstrap/data` folder, leaving the directory tree
looking like this:
```{r taf-bootstrap, echo=FALSE, results='asis', message=FALSE}
setwd(repoName)
taf.bootstrap()
setwd("..")
printDir()
```

## Preprocess data, write TAF data tables

Now that you have created the file structure and uploaded your data, its time to turn your attention to the `data.R` script. The  purpose of this script is to do some data
processing. This could be (among other things):

* calculating plus groups,
* taking averages of survey indices,
* calculating fillin values and removing outliers,
* calculating smoothed time series'

The `data.R` script should also write out input data into flat csv files, so they are readable and reviewable by others.

### Preprocessing

### Writing TAF tables
