---
output: github_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("R/utilities.R")
```

See also:
[Creating TAF datasets](Creating-TAF-datasets),
[Bib entries](Bib-entries),
[Example data records (bib entries)](Example-data-records).

This page was based on using the `icesTAF` package version ```r packageVersion("icesTAF")```
dated ```r packageDate("icesTAF")```.  

The resulting TAF analysis created from this example is 
on GitHub: [github.com/ices-taf-dev/wiki-example-1](https://github.com/ices-taf-dev/wiki-example-1),
if you dont want to build up the code yourself by following the example, you can skip to the end 
and run on your computer by downloading the complete code and running it like this:

```{r, eval = FALSE}
# get code
run_dir <- download.analysis("ices-taf-dev/wiki-example-1")
# view downloaded code
browseURL(run_dir)

# run analysis
run.analysis(run_dir)
# view result
browseURL(run_dir)
```


## In this guide

* [Creating an empty TAF project](#creating-an-empty-TAF-project)
* [Adding a local dataset](#adding-a-local-dataset)
* [Adding a local collection of data files](#adding-a-local-collection-of-data-files)
* [Adding a file from a URL](#adding-a-file-from-a-URL)
* [Adding data via a bootstrap script](#adding-data-via-a-bootstrap-script)

## Creating an empty TAF project

First we create an empty TAF project, and in this case we will call it `example-1`, and then moving our
working directory to this folder. We do this by running

```{r}
taf.skeleton("example-1")
setwd("example-1")
```

resulting in the following:
```{r, echo = FALSE, results="asis"}
tafdir <- "example-1"
# get skeleton path structure
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(tafdir, "/bootstrap/initial/data/"), paths)
printDir(paths)
```

## Adding a local dataset

First lets save or copy a file from somewhere else on our computer and put it into the 
`bootstrap\initial\data` folder. Remember, this is the place where you can bring in files that you
are unable to get from other online sources, such as web services and urls.

In this example we will use a dataset that comes shipped with R called `trees`, and save it as a
csv file.

```{r, echo=-1}
setwd(tafdir)
data(trees)
write.taf(trees, dir = "bootstrap/initial/data")
```

and now your project should look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(tafdir, "/bootstrap/initial/data/"), paths)
printDir(paths)
```

The way TAF works, is that only data in `bootstrap/data` are allowed to be used by the TAF scripts,
and the way that the `bootstrap/data` folder is populated is by creating entries in a file called
`DATA.bib`, and then running `taf.bootstrap()`.

We will create the `DATA.bib` file using the useful `draft.data()` function, and as we do so, will
add some useful information to document the dataset we are importing:

```{r, echo=-1}
setwd(tafdir)
draft.data(
  data.files = "trees.csv",
  data.scripts = NULL,
  originator = "Ryan, T. A., Joiner, B. L. and Ryan, B. F. (1976) The Minitab Student Handbook. Duxbury Press.",
  title = "Diameter, Height and Volume for Black Cherry Trees",
  file = TRUE,
  append = FALSE # create a new DATA.bib
)
```

after running
```{r, echo = -1}
setwd(tafdir)
taf.bootstrap()
```
your project should now look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

and you will have succesfully save, documented and imported (via running `taf.bootstrap()`) a
local dataset into your project.

## Adding a local collection of data files

In this example, we will add another dataset, but this time it will be a folder containing
several files.  You can create this yourself however you like, but the following code will create
an example for you

```{r, echo=-1}
setwd(tafdir)
data(trees)
data(cars)
# make the directory we want to write to
mkdir("bootstrap/initial/data/my-collection")
# save files there
write.taf(trees, dir = "bootstrap/initial/data/my-collection")
write.taf(cars, dir = "bootstrap/initial/data/my-collection")
```

and now your project should look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

Again we document this using `draft.data()` to add it to the `DATA.bib` file, but this time there are
two differences:

1. We are adding a new record, so we set `append = TRUE` as we want to add a record to an existing
list of records.
2. We are adding a folder, so we set `source = "folder"`

```{r, echo=-1}
setwd(tafdir)
draft.data(
  data.files = "my-collection",
  data.scripts = NULL,
  originator = "R datasets package",
  title = "Collection of R data",
  source = "folder",
  file = TRUE,
  append = TRUE # create a new DATA.bib
)
```

after running
```{r, echo = -1}
setwd(tafdir)
taf.bootstrap()
```
your project should now look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

and you will have succesfully save, documented and imported (via running `taf.bootstrap()`) a
local dataset, and a local collection of data files into your project.


## Adding a file from a URL

So far we have been importing local datasets and files, and so the `bootstrap/initial/data` folder
is identical to the `bootstrap/data` folder, and you may be wondering, 'what is the point?'.  An 
important purpose of this step is to add documentation to a dataset, to add provenance, so that 
every dataset in `bootstrap/data` has a coresponding record in `DATA.bib`. 

But this step does not just copy files from one place to the other. It can also fetch data and files
from other locations. In this example, we get a file from an URL, in this case, a raster file
of sea surface temperatures from the UK metoffice 
[www.metoffice.gov.uk/hadobs/hadsst4/](https://www.metoffice.gov.uk/hadobs/hadsst4/), and we can create
the entry in the `DATA.bib` file, again using `draft.data()`.

```{r, echo=-1}
setwd(tafdir)
draft.data(
  data.files = "HadSST.4.0.1.0_median.nc",
  data.scripts = NULL,
  originator = "UK MET office",
  title = "Met Office Hadley Centre observations datasets",
  year = 2022,
  source = "https://www.metoffice.gov.uk/hadobs/hadsst4/data/netcdf/HadSST.4.0.1.0_median.nc",
  file = TRUE,
  append = TRUE
)
```

after running
```{r, echo = -1}
setwd(tafdir)
taf.bootstrap()
```
your project should now look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

and you will have succesfully downloaded, documented and imported (via running `taf.bootstrap()`) a
dataset from a URL, and you will notice that the `bootstrap/data` folder now contains more than
what is in `bootstrap/initial/data`.

## Adding data via a bootstrap script

Sometimes it is not possible to download a dataset from a single url, and it requires multiple
steps to fetch and process data from an online source.  This is common with data accessed via
web services.  In this example we create a script where we write a short recipe of how to get
our data, and then register this in the `DATA.bib` file.  Firstly you need to write a script,
and for this example the following code will do that for you:

```{r, echo=-1}
setwd(tafdir)

cat('library(icesTAF)
library(sf)

download(
  "https://gis.ices.dk/shapefiles/OSPAR_Subregions.zip"
)

unzip("OSPAR_Subregions.zip")
unlink("OSPAR_Subregions.zip")

areas <- st_read("OSPAR_subregions_20160418_3857.shp")

# write as csv
st_write(
  areas, "ospar-areas.csv",
  layer_options = "GEOMETRY=AS_WKT"
)

unlink(dir(pattern = "OSPAR_subregions_20160418_3857"))
',
  file = "bootstrap/ospar-areas.R"
)
```

bootsrap scripts such as this one, goes in the `bootstrap` folder, 

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

and are decsribed in more detail in [Bib entries](Bib-entries). We then document it and add an 
entry to the `DATA.bib` however there are to differences to note:


```{r, echo=-1}
setwd(tafdir)
draft.data(
  data.files = NULL,
  data.scripts = "ospar-areas.R",
  originator = "OSPAR",
  title = "OSPAR areas",
  file = TRUE,
  append = TRUE
)
```

and after running
```{r, echo = -1}
setwd(tafdir)
taf.bootstrap()
```
your project should now look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

Now we have a folder with the same name as the script we created, and inside this is any files
created by the script.

```{r, echo = FALSE}
unlink("example-1", recursive = TRUE)
```
