---
output: github_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("R/utilities.R")
```

See also:
[Creating TAF datasets](Creating-TAF-datasets),
[Bib entries](Bib-entries),
[Example data records (bib entries)](Example-data-records).

This page was based on using the `icesTAF` package version ```r packageVersion("icesTAF")```
dated ```r packageDate("icesTAF")```.  

This example requires a package to be installed that is not on CRAN. Please ensure you have the
package `stockasessment` by running
```{r, eval = FALSE}
# Download and install stockassessment in R
install.packages("stockassessment", repos = "https://fishfollower.r-universe.dev")
```

The resulting TAF analysis created from this example is 
on GitHub: [github.com/ices-taf-dev/wiki-example-2](https://github.com/ices-taf-dev/wiki-example-2),
and can be run on your computer by:

```{r, eval = FALSE}
# get code
run_dir <- download.analysis("ices-taf-dev/wiki-example-2")
# view downloaded code
browseURL(run_dir)

# run analysis
run.analysis(run_dir)
# view result
browseURL(run_dir)
```


## In this guide

* [Creating an empty TAF project](#creating-an-empty-TAF-project)
* [Adding SAM input data via a bootstrap script](#adding-sam-input-data-via-a-bootstrap-script)

## Creating an empty TAF project

First we create an empty TAF project, and in this case we will call it `example-1`, and then moving our
working directory to this folder. We do this by running

```{r}
taf.skeleton("example-2")
setwd("example-2")
```

resulting in the following:
```{r, echo = FALSE, results="asis"}
tafdir <- "example-2"
# get skeleton path structure
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(tafdir, "/bootstrap/initial/data/"), paths)
printDir(paths)
```

## Adding SAM input data via a bootstrap script

Tso get the input data for a stock asseement run on 
[stockassessment.org/](https://stockassessment.org/), we can use a short script. It is provided here
as code that will create the script for you - normally you would create these scripts yourself.

```{r, echo=-1}
setwd(tafdir)

cat('library(stockassessment)

# download model from stockassessment.org
fit <- fitfromweb("WBCod_2021_cand01")

# save to model folder
save(fit, file = "fit.rData")

',
  file = "bootstrap/sam_fit.R"
)
```

bootsrap scripts such as this one, goes in the `bootstrap` folder:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

a second script to get the fitted model object is given below:

```{r, echo=-1}
setwd(tafdir)

cat('sam_assessment <- "WBCod_2021_cand01"

sam_dir <-
  paste0(
    "https://stockassessment.org/datadisk/stockassessment/userdirs/user3/",
    sam_assessment,
    "/data/"
  )

files <-
  paste0(
    c("cn", "cw", "dw", "lf", "lw", "mo", "nm", "pf", "pm", "survey", "sw"),
    ".dat"
  )

for (file in files) {
  download(paste0(sam_dir, file))
}

',
  file = "bootstrap/sam_data.R"
)
```

We can then add records for these scripts in the DATA.bib file using:

```{r, echo=-1}
setwd(tafdir)
draft.data(
  data.files = NULL,
  data.scripts = c("sam_data", "sam_fit"),
  originator = "WGBFAS",
  title = c("SAM input data for ...", "SAM fitted object for ..."),
  year = 2021,
  period = "1985-2020",
  file = TRUE,
  append = FALSE
)
```

resulting the project looking like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```


and after running
```{r, echo = -1}
setwd(tafdir)
taf.bootstrap()
```
your project should now look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

Now we have a folder with the same name as the script we created, and inside this is any files
created by the script.

```{r, echo = FALSE}
unlink("example-2", recursive = TRUE)
```
