---
output: github_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("R/utilities.R")
```

See also:
[Creating a TAF analysis](Creating-a-TAF-analysis),
[Bib entries](Bib-entries).

This page was based on using the `icesTAF` package version ```r packageVersion("icesTAF")```
dated ```r packageDate("icesTAF")```.

## In this guide

* [Using a script to download a zip file](#Using-a-script-to-download-a-zip-file)
* [Get the ICES Word template](#Get-the-ICES-Word-template)
* [Further `DATA.bib` entries](#Further-DATA.bib-entries)

## Using a script to download a zip file

We will start with an empty repository

```{r taf-skeleton-output, results='asis', echo=FALSE}
# make skeleton
taf.skeleton(repoName, force = TRUE)

# get skeleton path structure
paths <- dir(repoName, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(repoName, "/bootstrap/initial/data/"), paths)
printDir(paths)
```

Consider the following script that downloads a zip file containing ESRI
shapefiles of ICES areas, unzips it and deletes the zip file

```{r ices-areas, results='asis', echo=FALSE}
code <-
'filename <- "ICES_areas.zip"
# download and unzip
download(paste0("http://gis.ices.dk/shapefiles/", filename))
unzip(filename)
# delete zip file
unlink(filename)'
cat(formatRcode(code), sep = "\n")
```

This code can be used to download shapefiles for use in a TAF assessment by
specifying `script` as the `source` of the data.  This is done by creating
the following metadata record in `DATA.bib`

```{r draft-data, results='asis', echo=FALSE}
meta <-
  draft.data(
    data.files = "icesareas",
    originator = "ICES",
    title = "ICES Areas ESRI Shapefile",
    period = FALSE,
    source = "script"
  )
cat(formatRcode(meta), sep = "\n")
```

and an accompanying R script.  The R script must
have the same name as the 'key' field.  In this case the key is `icesareas` so
the script must be called `icesareas.R`.  The `DATA.bib` entry can be created
using the `r taflink("draft.data")` function

```{r draft-data2, echo = 2}
setwd(repoName)
  draft.data(
    data.files = "icesareas",
    originator = "ICES",
    title = "ICES Areas ESRI Shapefile",
    period = FALSE,
    source = "script",
    file = TRUE
  )
setwd("..")
```

After you have created the script in the bootstrap folder called `icesareas.R`
the directory structure should look like this:

```{r write-script, results='asis', echo=FALSE}
# save to bootstrap folder
setwd(repoName)
cat(code, file = "bootstrap/icesareas.R", append = FALSE)
setwd("..")
printDir()
```

Now we can run  `r taflink("taf.bootstrap")` to process the script to download the
process the zip file and now the directory structure will look like this

```{r process-script, echo=2, results='asis', message=FALSE}
setwd(repoName)
taf.bootstrap()
setwd("..")
printDir()
```

## Get the ICES Word template

```{r setup2, echo=FALSE, message=FALSE, warning=FALSE}
source("R/utilities.R")
```

We will start with an empty repository

```{r taf-skeleton-output2, results='asis', echo=FALSE}
# make skeleton
taf.skeleton(repoName, force = TRUE)

# get skeleton path structure
paths <- dir(repoName, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(repoName, "/bootstrap/initial/data/"), paths)
printDir(paths)
```

To download this file for use in an automated report for a
TAF assessment, specify the location of the file as the `source`
of the data record.  The file is held in the
[ices-taf/doc](https://github.com/ices-taf/doc) repository
and is called reportTemplate.docx. This is done by creating the following
meta-data record in `DATA.bib`

```{r draft-data3, results='asis', echo=FALSE}
meta <-
  draft.data(
    data.files = "reportTemplate.docx",
    originator = "ICES",
    title = "ICES TAF Word template for report automation",
    period = FALSE,
    source = "https://github.com/ices-taf/doc/raw/master/reportTemplate.docx"
  )
cat(formatRcode(meta), sep = "\n")
```

The `DATA.bib` entry can be created using the `r taflink("draft.data")`
function

```{r draft-data4, echo = 2}
setwd(repoName)
  draft.data(
    data.files = "reportTemplate.docx",
    originator = "ICES",
    title = "ICES TAF Word template for report automation",
    period = FALSE,
    source = "https://github.com/ices-taf/doc/raw/master/reportTemplate.docx",
    file = TRUE
  )
setwd("..")
```

The directory structure should now look like this:

```{r print-dir, results='asis', echo=FALSE}
printDir()
```

Now we can run  `r taflink("taf.bootstrap")` to process the script to
download the Word template file and now the directory structure will look like this

```{r process-word-template, echo=2, results='asis', message=FALSE}
setwd(repoName)
taf.bootstrap()
setwd("..")
printDir()
```

## Further `DATA.bib` entries

#### ICES Statistical Rectangles mapped to Ecoregions

`DATA.bib` entry:

```r
@Misc{ICES_StatRec_mapto_Ecoregions,
  originator = {DTU Aqua},
  year       = {2019},
  title      = {ICES Stat rec ESRI Shapefile},
  url        = {https://gis.ices.dk/geonetwork/srv/metadata/81f68a99-9b91-4762-80d3-31c069731f44},
  source     = {script},
}
```

with the R script: `ICES_StatRec_mapto_Ecoregions.R`

```r
filename <- "ICES_StatRec_mapto_Ecoregions.zip"
# download and unzip
download(paste0("http://gis.ices.dk/shapefiles/", filename))
unzip(filename)
# delete zip file
unlink(filename)
```

### ICES Ecoregion

`DATA.bib` entry:

```r
@Misc{ICES_ecoregions,
  originator = {ICES},
  year       = {2019},
  title      = {ICES Ecoregion ESRI Shapefile},
  url        = {https://gis.ices.dk/geonetwork/srv/metadata/4745e824-a612-4a1f-bc56-b540772166eb},
  source     = {script},
}
```

with the R script: `ICES_ecoregions.R`

```r
filename <- "ICES_ecoregions.zip"
# download and unzip
download(paste0("http://gis.ices.dk/shapefiles/", filename))
unzip(filename)
# delete zip file
unlink(filename)
```

### Species lookup table from ICES SD database

`DATA.bib` entry:

```r
@Misc{ICES_SD_species_lookup,
  originator = {ICES},
  year       = {2019},
  title      = {ICES Fisheries Guild lookup table},
  url        = {https://gis.ices.dk/geonetwork/srv/metadata/30541cf4-0236-437f-9757-596c5f793cff},
  source     = {script},
}
```

with the R script: `ICES_SD_species_lookup.R`
```r
library(icesSD)
library(magrittr)

sid <- getSD()

# get lookup table for species, common name and Fisheries guild from SID
species_lookup <-
  sid %>%
  filter(ActiveYear > 2018) %>%
  select(
    SpeciesScientificName,
    SpeciesCommonName,
    FisheriesGuild
  ) %>%
  mutate(
    FisheriesGuild = tolower(FisheriesGuild)
  ) %>%
  filter(
    !is.na(FisheriesGuild) &
    !is.na(SpeciesScientificName)
  ) %>%
  unique()

write.taf(species_lookup, quote = TRUE)
```
