---
output: github_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("R/utilities.R")
```

See also:
[Getting set up](Getting-set-up), 
[Running a TAF analysis](Running-a-TAF-analysis),
[Creating TAF datasets](Creating-TAF-datasets).

This page was based on using the `icesTAF` package version ```r packageVersion("icesTAF")```
dated ```r packageDate("icesTAF")```.  

This example requires a package to be installed that is not on CRAN. Please ensure you have the
package `stockasessment` by running
```{r, eval = FALSE}
# Download and install stockassessment in R
install.packages("stockassessment", repos = "https://fishfollower.r-universe.dev")
```

The resulting TAF analysis created from this example is 
on GitHub: [github.com/ices-taf-dev/wiki-example-3](https://github.com/ices-taf-dev/wiki-example-3),
if you dont want to build up the code yourself by following the example, you can skip to the end 
and run on your computer by downloading the complete code and running it like this:

```{r, eval = FALSE}
# get code
run_dir <- download.analysis("ices-taf-dev/wiki-example-3")
# view downloaded code
browseURL(run_dir)

# run analysis
run.analysis(run_dir)
# view result
browseURL(run_dir)
```


## In this guide

* [Creating an empty TAF project](#creating-an-empty-TAF-project)
* [Preprocess the data](#preprocess-the-data)
* [Running a model](#running-a-model)
* [Writing TAF tables](#writing-taf-tables)
* [Formatted output for reporting](#formatted-output-for-reporting)

## Creating an empty TAF project

First we create an empty TAF project set up with a bootstrap section for a stock assessment
created on stock [stockassessment.org/](https://stockassessment.org/), and in this case we will 
call it `example-3`, and then moving our working directory to this folder. We do this by running

```{r}
taf.skeleton.sa.org("example-3", "WBCod_2021_cand01")
setwd("example-3")
```

resulting in the following:
```{r, echo = FALSE, results="asis"}
tafdir <- "example-3"
# get skeleton path structure
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(tafdir, "/bootstrap/initial/data/"), paths)
printDir(paths)
```

and after running
```{r, echo = -1}
setwd(tafdir)
taf.bootstrap()
```
your project should now look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

Now we have a folder with the same name as the script we created, and inside this is any files
created by the script.

## Preprocess the data

Now that you have created the file structure and uploaded your data, it's time to turn your attention to the `data.R` script. The  purpose of this script is to do some data
processing. This could be (among other things):

* calculating a plus group,
* taking averages of survey indices,
* calculating fill-in values and removing outliers,
* calculating smoothed time series

The `data.R` script should also write out input data into flat `.csv` files, so they are readable and reviewable by others.

## Running a model

## Writing TAF tables

## Formatted output for reporting 

```{r, echo = FALSE}
unlink("example-3", recursive = TRUE)
```
