---
output: github_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("R/utilities.R")
```

See also:
[Getting set up](Getting-set-up), 
[Running a TAF analysis](Running-a-TAF-analysis),
[Creating TAF datasets](Creating-TAF-datasets).

This page was based on using the `icesTAF` package version ```r packageVersion("icesTAF")```
dated ```r packageDate("icesTAF")```.  

This example requires a package to be installed that is not on CRAN. Please ensure you have the
package `stockasessment` by running
```{r, eval = FALSE}
# Download and install stockassessment in R
install.packages("stockassessment", repos = "https://fishfollower.r-universe.dev")
```

The resulting TAF analysis created from this example is 
on GitHub: [github.com/ices-taf-dev/wiki-example-stockassessment](https://github.com/ices-taf-dev/wiki-example-stockassessment),
if you dont want to build up the code yourself by following the example, you can skip to the end 
and run on your computer by downloading the complete code and running it like this:

```{r, eval = FALSE}
# get code
run_dir <- download.analysis("ices-taf-dev/wiki-example-stockassessment")
# view downloaded code
browseURL(run_dir)

# run analysis
run.analysis(run_dir)
# view result
browseURL(run_dir)
```


## In this guide

* [Creating an empty stockassessment.org TAF project](#creating-an-empty-stockassessment.org-TAF-project)
* [Preprocess the data](#preprocess-the-data)
* [Running a model](#running-a-model)
* [Writing TAF tables](#writing-taf-tables)
* [Formatted output for reporting](#formatted-output-for-reporting)

## Creating an empty stockassessment.org TAF project

First we create an empty TAF project set up with a bootstrap section for a stock assessment
created on stock [stockassessment.org/](https://stockassessment.org/), and in this case we will 
call it `example-4` and choose the run called: `WBCod_2021_cand01`, and then moving our working 
directory to this folder. We do this by running

```{r}
taf.skeleton.sa.org("example-4", "WBCod_2021_cand01")
setwd("example-4")
```

resulting in the following:
```{r, echo = FALSE, results="asis"}
run_dir <- download.analysis("ices-taf-dev/wiki-example-stockassessment")
tafdir <- "example-4"
# get skeleton path structure
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(tafdir, "/bootstrap/initial/data/"), paths)
printDir(paths)
```

and after running
```{r, echo = -1}
setwd(tafdir)
taf.bootstrap()
```
your project should now look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```

Now we have a input data and the fitted model object that we can use to create some standard TAF
files.

## Preprocess the data

Now that you have created the file structure and uploaded your data, it's time to turn your attention to the `data.R` script. The  purpose of this script is to do some data
processing. This could be (among other things):

* calculating a plus group,
* taking averages of survey indices,
* calculating fill-in values and removing outliers,
* calculating smoothed time series

The `data.R` script should also write out input data into flat `.csv` files, so they are readable and reviewable by others.

In this section we will build the script up in sections.  First we need to load the libraries and get the data required 
for the script. It is good practice to write your analysis in several scripts with each script being as independent as 
possible. This will allow you to return to your analysis and work on a specific section, and not worry too much about
whether you have loaded the correct libraries and data.  So the first step is always to load libraries, then load the data 
required.

It is also considered good practice to create (or ensure) that the output directory for the TAF section you are working in
exists. Hence the start of your script will be structured like this:

```{r, eval = FALSE}
## Preprocess data, write TAF data tables

## Before:
## After:

## load libraries
library(icesTAF)
library(stockassessment)

# ensure directory
mkdir("data")

#  Read underlying data from bootstrap/data
# ...
```

For the example we are working with here this becomes

```{r, eval = FALSE}
## Preprocess data, write TAF data tables

## Before:
## After:

## load libraries
library(icesTAF)
library(stockassessment)

# ensure directory
mkdir("data")

#  Read underlying data from bootstrap/data

#  ## Catch-numbers-at-age ##
catage <- read.ices(taf.data.path("sam_data/cn.dat"))

#  ## Catch-weight-at-age ##
wcatch <- read.ices(taf.data.path("sam_data/cw.dat"))
wdiscards <- read.ices(taf.data.path("sam_data/dw.dat"))
wlandings <- read.ices(taf.data.path("sam_data/lw.dat"))

#  ## Natural-mortality ##
natmort <- read.ices(taf.data.path("sam_data/nm.dat"))

#  ## Proportion of F before spawning ##
propf <- read.ices(taf.data.path("sam_data/pf.dat"))

#  ## Proportion of M before spawning ##
propm <- read.ices(taf.data.path("sam_data/pm.dat"))

#  ## Stock-weight-at-age ##
wstock <- read.ices(taf.data.path("sam_data/sw.dat"))

# Landing fraction in catch at age
landfrac <- read.ices(taf.data.path("sam_data/lf.dat"))
```

The next step is to do any preprocessing or calculations that you may need to do.  In this simple
example we don't need to do much

```{r, eval=FALSE}
## 2 Preprocess data

# landings
latage <- catage * landfrac[, -1]
datage <- catage * (1 - landfrac[, -1])
```

And finally when we are done, it is recommended that you write out your input data in a human (easily) readable
form.  CSV is always a good option as it is easily read into R and Excel etc.

```{r, eval=FALSE}
## 3 Write TAF tables to data directory
write.taf(
  c(
    "catage", "latage", "datage", "wstock", "wcatch",
    "wdiscards", "wlandings", "natmort", "propf", "propm",
    "landfrac"),
  dir = "data"
)
```

And this concludes the data script. To test the script you can run `sourceTAF("data")

```{r, echo=FALSE}
cp(file.path(run_dir, "data.R"), file.path(tafdir))
```

```{r, echo=-1}
setwd(tafdir)
sourceTAF("data")
```
your project should now look like this:

```{r, echo = FALSE, results="asis"}
paths <- dir(tafdir, full.names = TRUE, recursive = TRUE)
printDir(paths)
```


## Running a model

## Writing TAF tables

## Formatted output for reporting 

```{r, echo = FALSE}
unlink("example-4", recursive = TRUE)
```
