---
output: github_document
---

  
```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(icesTAF)
library(data.tree)
repoName <- "cod.27.47d20"

if (dir.exists(repoName)) {
  unlink(repoName, recursive = TRUE)
}

formatRcode <- function(x) {
  c("\`\`\`r",
    x,
    "\`\`\`")
}

printDir <- function() {
  # get skeleton path structure
  paths <- dir(repoName, full.names = TRUE, recursive = TRUE)

  # make a data.tree and print it
  tree <- as.Node(data.frame(pathString = paths))
  res <- capture.output(print(as.data.frame(tree), row.names = FALSE))[-1]
  cat(formatRcode(res), sep = "\n")
}

taflink <- function(x) {
  sprintf("[`%1$s`](https://rdrr.io/cran/icesTAF/man/%1$s.html)", x)
}
```

This page was based on using the `icesTAF` package version ```r packageVersion("icesTAF")```
dated ```r packageDate("icesTAF")```.

## In this guide

* [Using a script](#Using-a-script)


We will start with an empty repository
```{r taf-skeleton-output, results='asis', echo=FALSE}
# make skeleton
taf.skeleton(repoName, force = TRUE)

# get skeleton path structure
paths <- dir(repoName, full.names = TRUE, recursive = TRUE)
paths <- c(paste0(repoName, "/bootstrap/initial/data/"), paths)

# make a data.tree and print it
tree <- as.Node(data.frame(pathString = paths))
res <- capture.output(print(as.data.frame(tree), row.names = FALSE))[-1]
cat(formatRcode(res), sep = "\n")
```


## Using a script

Consider the following script that downloads a zip file containing ESRI
shapefiles of ICES areas, unzips it and deletes the zip file
```{r ices-areas, results='asis', echo=FALSE}
code <- 
'filename <- "ICES_areas.zip"
# download and unzip
download(paste0("http://gis.ices.dk/shapefiles/", filename))
unzip(filename)
# delete zip file
unlink(filename)'
cat(formatRcode(code), sep = "\n")
```

This code can be used to download shapefiles for use in a TAF assessment by
specifying `script` as the `source` of the data.  This is done by creating
the following meta-data record in `DATA.bib`

```{r draft-data, results='asis', echo=FALSE}
meta <- 
  draft.data(data.files = "icesareas",
             originator = "ICES", 
             title = "Catch data for cod.27.347d", 
             period = "2015-2018",
             source = "script")
cat(formatRcode(meta), sep = "\n")
```

and an accompaning R script.  The R script must
have the same name as the 'key' feild.  In this case the key is `icesareas` so
the script must be called `icesareas.R`.  The `DATA.bib` entry can be created
using the `r taflink("draft.data")` function


```{r draft-data2, echo = 2}
setwd(repoName)
draft.data(data.files = "icesareas",
           originator = "ICES", 
           title = "Catch data for cod.27.347d", 
           period = "2015-2018",
           source = "script",
           file = "bootstrap/DATA.bib")
setwd("..")
```

after you have created the script in the bootstrap folder called `icesareas.R`
the directory structure should look like this:

```{r write-script, results='asis', echo=FALSE}
# save to bootstrap folder
setwd(repoName)
cat(code, file = "bootstrap/icesareas.R", append = FALSE)
setwd("..")
printDir()
```

Now we can run  `r taflink("taf.bootstrap")` to process the script to download the 
process the zip file and now the directory structyre will look like this

```{r process-script, echo=2, results='asis', message=FALSE}
setwd(repoName)
taf.bootstrap()
setwd("..")
printDir()
```

