---
output: github_document
---

## In this guide

- [Software to install](#software-to-install)
- [R packages to install](#r-packages-to-install)
- [Additional software to install](#additional-software-to-install)

## Software to install

The easiest way to set up your system is to install:

* R: [cran.r-project.org/](https://cran.r-project.org/)
* Rstudio: [www.rstudio.com/products/rstudio/download](https://www.rstudio.com/products/rstudio/download/#download)

## R packages to install

Once you have installed R (and RStudio) you can begin to install some R packages.  For TAF to work
you need only one package: `icesTAF`.  You install this by running the line:

```{r, eval = FALSE}
install.packages("icesTAF")
```

However, to run the code in this wiki, you need the development version, which can be found here:
```{r, eval = FALSE}
install.packages("icesTAF", repos = "https://ices-tools-prod.r-universe.dev")
```



## Additional software to install

If you wish to also use GitHub to version control your code, then you also need to install:
* Git: [https://git-scm.com/downloads](https://git-scm.com/downloads)

And finally if you are using Windows, and you wish to install some special R packages from GitHub
you will need to install:
* Rtools: [cran.r-project.org/bin/windows/Rtools](https://cran.r-project.org/bin/windows/Rtools/)
