---
output: github_document
---

# Very quick guide to developing R packages

time to work through: ~ 1hr

## reading list:

* main reference
  - https://r-pkgs.org/

* documenting R packages using roxygen2
  - https://CRAN.R-project.org/package=roxygen2
  - https://cran.r-project.org/web/packages/roxygen2/vignettes/roxygen2.html

* some example packages:
  - https://github.com/ices-tools-prod

```{r, setup}
library(devtools)
library(usethis)
library(git2r)
# install_github("fledge-iot/fledge")
library(fledge)
unlink("myTest", recursive = TRUE)
```


### minimal package

```{r, eval = FALSE}
create_package("myTest", fields = list(Date = Sys.Date()))
git2r::init()
git2r::add(path = "*")
git2r::commit(message = "initial create")
```

### Add a license

ICES DIG is proposing to go down an open source license root.  MIT is
designed to be an interuseable license, so that your software can be
reused in many other projects, while still retaining a reference back
to you as copyright holder

```{r, eval = FALSE}
copyrightHolders <-
  c(
    "Colin Millar",
    "Where I work maybe",
    "Other authors"
  )
use_mit_license(copyrightHolders)
git2r::add(path = "*")
git2r::commit(message = "add license")
```

### develop, document and check cycle

* create a new file, with the same name as the function
* code up the function
* add documentation
* create documentation
* check

```{r, eval = FALSE}
cat(
"
#' Capitalise the first letter of each word in a string of words
#'
#' @export
capitalise <- function(string) {
  words <- strsplit(string, ' ')[[1]]
  Words <-
    paste0(
      toupper(substring(words, 1, 1)),
      substring(words, 2)
    )
  paste(Words, collapse = ' ')
}
",
file = "R/capitalise.R")
```

document and check
```{r, eval = FALSE}
devtools::document()
devtools::check()
```

fix the warnings: undocumented arguments

```{r, eval = FALSE}
cat(
"
#' Capitalise the first letter of each word in a string of words
#'
#' @param string a character string with words seperated by a space
#'
#' @export
capitalise <- function(string) {
  words <- strsplit(string, ' ')[[1]]
  Words <-
    paste0(
      toupper(substring(words, 1, 1)),
      substring(words, 2)
    )
  paste(Words, collapse = ' ')
}
",
file = "R/capitalise.R")
```

document and check
```{r, eval = FALSE}
devtools::document()
devtools::check()
```

all OK now :)
commit news worthy item (put a star (*) in front of message)
```{r, eval = FALSE}
git2r::add(path = "*")
git2r::commit(message = "* added capitalise() function")
```

### add a data set

```{r, eval = FALSE}
usethis::use_data_raw()
# remove DATASET.R created by use_data_raw()
# useful to look at though
unlink("data-raw/DATASET.R")
cat("
salmon <- read.csv(
  'https://data.marine.gov.scot/sites/default/files//Girnock_Baddoch_Emigrant_Numbers_by_Year_of_Emigration_2020.csv'
)
names(salmon) <- tolower(names(salmon))
usethis::use_data(salmon, overwrite = TRUE)
",
  file = "data-raw/salmon.R"
)
# source this - but in practice you just run from your editor
source("data-raw/salmon.R")
```
check should fail
```{r, eval = FALSE}
devtools::check()
```

now document data set:
```{r, eval = FALSE}
cat("
#' Girnock and Baddoch: Emigrant Numbers by Year of Emigration.
#'
#' The Girnock and Baddoch catchments are long-term monitoring sites
#' where the population dynamics of spring salmon have been assessed
#' since 1966 and 1988 respectively. These sites include a pair of
#' fixed traps designed to catch returning adult salmon and descending
#' juvenile emigrants. The descending traps are whole river traps based
#' on a modified Wolf trap, and operate on permanent weirs. They are
#' routinely operated in both autumn and spring, providing counts of
#' emigrant salmon. Trap protocols have been similar, although not
#' identical, over this long period.
#'
#' @format A data frame with 54 rows and 5 variables:
#' \\describe{
#'   \\item{year}{year of counts}
#'   \\item{girnock.smolts}{counts of smolts at the Girnock trap}
#'   \\item{girnock.parr}{counts of parr at the Girnock trap}
#'   \\item{baddoch.smolts}{counts of smolts at the Baddoch trap}
#'   \\item{baddoch.parr}{counts of parr at the Baddoch trap}
#' }
#' @source \\url{https://data.marine.gov.scot/dataset/girnock-and-baddoch-emigrant-numbers-year-emigration}
'salmon'
",
  file = "R/data.R"
)
```

document and check
```{r, eval = FALSE}
devtools::document()
devtools::check()
```

checks pass - make sure items match with column names etc.

commit news worthy item
```{r, eval = FALSE}
git2r::add(
  path = c(
    "data/salmon.rda", "data-raw/salmon.R",
    "R/data.R", "man/salmon.Rd",
    ".Rbuildignore",
    "DESCRIPTION"
  )
)
git2r::commit(message = "* added salmon dataset")
```

### bump version (if you are maintainer)

this collects all news worth commits and adds them to the news file

```{r, eval = FALSE}
fledge::bump_version()
```



### local testing

```{r, eval = FALSE}
# development
devtools::load_all()
# check package as a user
devtools::install()
library("rdbesTest")
?salmon
?capitalise
```

### testing - simple

add an example to the function capitalise()

```{r, eval = FALSE}
cat(
"
#' Capitalise the first letter of each word in a string of words
#'
#' @param string a character string with words seperated by a space
#'
#' @examples
#' capitalise('hello world')
#'
#' @export
capitalise <- function(string) {
  words <- strsplit(string, ' ')[[1]]
  Words <-
    paste0(
      toupper(substring(words, 1, 1)),
      substring(words, 2)
    )
  paste(Words, collapse = ' ')
}
",
file = "R/capitalise.R")
```

document and check
```{r, eval = FALSE}
devtools::document()
devtools::check()
# check package as a user
devtools::install()
library("myTest")
?capitalise
example(capitalise)
# the ultimate is pushing to github and installing via:
# devtools::install_github("mygithubaccount/myTest")
```

commit non-news worthy item
```{r, eval = FALSE}
git2r::add(path = "*")
git2r::commit(message = "added examples to capitalise")
```

### testing

use testhat package

```{r, eval = FALSE}
# set up structure for testing
usethis::use_testthat()
usethis::use_test("capitalise")
# overwrite capitalise.R script - it is useful to read it before
# we overwrite though.
cat("# test strings
hello_world <- 'hello world'
test_that('a single string is returned', {
  testthat::expect_length(capitalise(hello_world), 1)
})
",
  file = "tests/testthat/test-capitalise.R"
)
```

run tests

```{r, eval = FALSE}
devtools::test()
```

commit news worthy item
```{r, eval = FALSE}
git2r::add(
  path = c(
    "tests/*",
    "DESCRIPTION"
  )
)
git2r::commit(message = "* added tests for capitalise")
```

### review commit history

```{r, eval = FALSE}
git2r::commits()
```

### ReadMe markdown file

still to do!